\papertitle

### 2025-05-05 paper/main.tex
add data availability statement
add conflict of interest statement

% !TeX program = pdflatex
% Minimally-Congested Travel Time Prediction from Sparse Open Data
% Authors: Geoff Boeing and Yuquan Zhou
% Web: https://geoffboeing.com/
% Repo: https://github.com/gboeing/travel-time-prediction

\newcommand{\myname}{Geoff Boeing and Yuquan Zhou}
\newcommand{\myemail}{boeing@usc.edu}
\newcommand{\myaffiliation}{University of Southern California}
\newcommand{\paperdate}{2025}
\newcommand{\papertitle}{Minimally-Congested Travel Time Prediction from Sparse Open Data}
\newcommand{\papercitation}{Boeing, G. and Y. Zhou. \paperdate. \papertitle. Under review at \emph{Journal Name}.}
\newcommand{\paperkeywords}{Urban Planning, Transportation, Data Science}

@@

\title{\papertitle}
\author[]{Redacted for review}%{\myname}
\affil[]{Redacted for review}%{\myaffiliation}
\date{}%\paperdate}

@@

Travel time prediction is central to accessibility analyses, sustainable transportation infrastructure provision, and active transportation interventions. However, calculating accurate travel times, especially for driving, requires either extensive technical capacity and bespoke data, or sources like the Google Maps API that quickly become prohibitively expensive to analyze thousands or millions of trips necessary for metropolitan-scale planning. These obstacles make it difficult for researchers, practitioners, and community advocates to develop an evidence base for more sustainable transportation and equitable accessibility. This article introduces a free, open-source minimally-congested driving time prediction model with minimal cost, data, and computational requirements. It trains and tests this model using the Los Angeles, California urban area as a case study by calculating naïve travel times from open data then developing a random forest model to predict true travel times as a function of those naïve times plus open data on turns and traffic controls. Validation shows that this method produces accurate, free, and simple travel time predictions, offering a superior middle-ground technique for urban planners.

@@

Travel time prediction underpins many transport planning processes and research problems. It is key to understanding urban accessibility, transport mode and route choice, and location decisions \citep{zhang2015gradient,jenelius2013travel}. Predicting congested travel times (i.e., when traffic congestion impedes flows) requires large volumes of real-time data on traffic conditions \citep[e.g.,][]{hou2018network}. Such data are proprietary, expensive, and beyond the reach of many planning practitioners and scholars \citep{carrion2012value,wang2011estimating}. This is usually not a problem for well-resourced organizations with highly-skilled labor---however, technical expertise, complex data and software requirements, and high costs pose insurmountable obstacles for many urban planning researchers, practitioners, and community advocates \citep{giles2022creating}.

Although typical \enquote{naïve} travel time prediction methods (which minimize travel distance or street segment traversal time) provide easy solutions, they ignore delays from turns and traffic controls and thus usually under-predict real-world driving travel time \citep{ludwig2023traffic,salonen2013modelling,yiannakoulias2013estimating}. However, as we will argue, minimally-congested travel time prediction offers a better middle ground of more-accurate predictions than naïve travel times do---yet much lower data and computing requirements than real-time traffic-aware congested travel time prediction.

To address this need for a middle ground, this article contributes a novel accurate, easy-to-use, and free method of minimally-congested driving travel time prediction. Our results show a substantial improvement over traditional, naïve methods of travel time prediction common in the urban planning literature. As we shall show in this article, across the Los Angeles urban area, our average trip's travel time prediction differs from Google's \enquote{gold standard} prediction by only 0.45 seconds. In contrast, simple---but common---naïve travel time prediction minimizing max-speed edge traversal times under-predicts by 183.55 seconds on average.

The rest of this article is organized as follows. First it reviews the state of the art in travel time prediction and then considers the less-sophisticated methods that are standard in the current planning literature. Next it describes our data sources, models, and validation techniques. Then it presents our findings, highlighting both the feasibility and quality of our predictions in relation to simple naïve predictions or difficult expensive predictions. Finally, it concludes with a discussion of implications for research and practice.

@@

Modern travel time prediction techniques tend to fall into one of two camps: (1) sophisticated, accurate, but difficult to execute; or (2) naïve, less-accurate, but simple to execute. The former is the state-of-the-art and has been the subject of much work in the recent computer science and engineering literatures. The latter tends to appear often in urban planning research and practice. Here we summarize the recent advances and current standards in these fields.

@@

A better way would be to seek a middle ground between expensive, complicated, point-to-point empirical travel times and overly-simple naïve predictions. Such free empirical travel time predictions may not reflect congested travel times (which vary drastically throughout the day and week), but would provide a good prediction of minimally-congested travel times. This middle-ground method should meet three criteria. First, it should only use free software and free data, avoiding expensive API queries or bespoke GPS data collection. Second, it should be a simple and accessible method for urban planners without advanced technical skills. Third, it should provide a substantial improvement in accuracy or spatial resolution over naïve predictions such as minimizing Euclidean distance, network distance, street edge traversal time, and exsiting TAZ-to-TAZ travel times, and approach the accuracy of the state-of-art models. Meeting these three criteria, a middle ground should be inexpensive, easy-to-use, and reasonably accurate.

@@

This article proposes a novel open-source, reusable, generalizable method\endnote{Our open-source tool is hosted at [REDACTED] for others to freely reuse.} to solve this problem. In short, it collects free open data plus a small amount of free data from the Google Maps API to train a local model to predict minimally-congested driving travel times with sparse data and simple computing hardware. \autoref{fig:workflow} shows a detailed workflow.

@@

After tuning, the four model training algorithms perform similarly across most---but not all---of the accuracy indicators (\autoref{tab:validation_results}). They fall roughly in a similar range for MAPE (7.8--9.0\%), MAE (71.8--79.7 seconds), MSE (12,140--13,517 square seconds), APR (0.99--1.01), and $R^2$ (0.93), demonstrating the robustness of our model to algorithmic particulars. However, the random forest and decision tree models exhibit much lower absolute systematic prediction bias ($\delta$ = 0.45 and 0.44 seconds, respectively) than the gradient boosting and AdaBoost models do (-19.3 and -9.7 seconds, respectively). In other words, all these models have comparable mean errors---but the average random forest and decision tree predictions are statistically insignificantly different from the average true travel time, whereas gradient boosting and AdaBoost significantly under-predict it.

Therefore, we select the random forest model as our \enquote{final} travel time prediction model because it predicts better overall (than the decision tree model) and better avoids systematic under-prediction (than the gradient boosting and AdaBoost models). This final model trains an ensemble of decision trees to predict unobserved response values by essentially dividing the data into boxes then making predictions based on the means of those boxes. Its optimized hyperparameterization uses 400 decision trees, random sampling with replacement in each decision tree, a maximum decision tree depth of 10, use of all available input explanatory features at each split, a requirement of at least two samples to split at a decision node, and the default settings that require at least one sample at a leaf node and apply equal sample weighting.

Finally, we check this final model for overfitting by conducting another 5-fold cross-validation across the whole sample using the tuned hyperparameters. The resulting five MAE values (75.3, 74.5, 73.5, 73.3, 74.6) are all very close to each other, indicating that the model is not overfitted and its hyperparameters are not overly specific to the training set.

@@

\begin{table*}[tb!]
    \centering\small
    \caption{Out-of-sample prediction accuracy of our chosen model (random forest, bold) versus the initial naïve travel time model and the discarded alternative models, all validated against the corresponding true travel times. $n$ = 8,272 for each. The six accuracy indicators are (1) MAPE \%, (2) MAE in seconds, (3) MSE in square seconds, (4) difference-in-means ($\delta$) in seconds and its corresponding $t$-test's $p$-value, (5) APR, and (6) $R^2$. See methods for definitions.}\label{tab:validation_results}
    \begin{tabular}{lrrrrrrr}
        \toprule
        Model               & MAPE  & MAE    & MSE      & $\delta$ & $p$  & APR  & $R^2$ \\
        \midrule
        Initial naïve       & 21.13 & 183.55 & 48155.43 &  -182.71 & <0.01 & 0.79 & 0.74  \\
        \textbf{Random forest}       &  \textbf{8.43} &  \textbf{75.24} & \textbf{12140.14} &     \textbf{0.45} &  \textbf{0.71} & \textbf{1.01} & \textbf{0.93}  \\
        Gradient boosting   &  7.84 &  71.81 & 12572.83 &   -19.30 & <0.01 & 0.99 & 0.93  \\
        Decision trees      &  8.96 &  79.74 & 13517.03 &     0.44 &  0.73 & 1.01 & 0.93  \\
        AdaBoost            &  8.21 &  74.05 & 12416.08 &    -9.73 & <0.01 & 1.00 & 0.93  \\
        \bottomrule
    \end{tabular}
\end{table*}

Compared to the initial naïve travel time calculation, our travel time prediction model exhibits substantial out-of-sample improvement (\autoref{tab:validation_results}). Our model produces a MAPE of 8.4\%, in line with the \textasciitilde10\% MAPEs seen in the state-of-the-art travel time prediction literature, but without their extensive and expensive input data requirements. In comparison, our initial naïve travel time calculation produces a much worse MAPE of 21.1\%. Whereas the naïve travel time's MAE is 183.6 seconds, our model's predicted travel times' MAE is just 75.2 seconds---an improvement by a factor of 2.4. Similarly for MSE, our model improved by a factor of 4.0.

The $t$-test reveals significant differences between the initial naïve travel time calculations and the true travel times: the $\delta$ of -183.6 seconds corresponds to $p$ < 0.01. In other words, the naïve travel time under-predicts true travel time by over 3 minutes. However, the $t$-test reveals insignificant differences between our model's out-of-sample predictions and the true travel times: the $\delta$ of 0.45 seconds corresponds to a $p$ value of 0.71. That is, our prediction model over-predicts by less than half a second on average, which is statistically insignificantly different from zero. If we reconsider the aforementioned MAEs in light of these differences-in-means, we can see that the initial naïve travel time calculations significantly and consistently under-predict travel time, but our prediction model shows no such directional bias: its random error averages out between (smaller absolute) over-and under-prediction.

@@

The planning literature shows that planners often use driving travel time prediction techniques far below the state-of-the-art due to cost, data, and technical capacity constraints. This study investigated if a better middle-ground travel time prediction method was feasible. It identified three criteria for success: (1) it should use free software and data and avoid bespoke sensor data collection; (2) it should be easy and accessible to use; (3) it should offer better accuracy than naïve predictions (e.g., minimizing Euclidean distance, network distance, street edge traversal time) and better spatial resolution than zone-to-zone travel times.

@@

This middle ground is not necessarily for computer scientists, high-end labs with top-of-the-line computing hardware, or well-resourced agencies with extensive data budgets. Rather, it empowers the scholars and practitioners who need it the most---those working on the frontlines of urban planning and policymaking. The literature shows that these scholars and practitioners often rely on simple but inaccurate routing models---such as minimizing Euclidean distance, network distance, or edge traversal time---when they lack the technical expertise, computing equipment, or extensive data needed to implement the literature's cutting-edge models. However, this study demonstrates how these naïve methods systematically under-predict real-world travel times. Our model addresses this problem by incorporating sparse open data on traffic control elements and turns into the prediction to achieve an accuracy in line with state-of-the-art methods---but without their data needs or computing requirements. In other words, our model does not replace the state-of-the-art congested-traffic methods for those with extensive resources available to implement them, but rather offers far more accurate predictions for urban planners who would otherwise fall back on simpler naïve models. Inaccurate travel times skew empirical understandings and mislead planning interventions. Our model represents an important middle-ground contribution for urban planning scholarship and evidence-informed practitioner interventions.

As demonstrated, this model supports high-resolution point-to-point travel time predictions across a large metropolitan area. Our implementation focused on Los Angeles but is not inherently tied to it, and future work should emphasize its generalizability in cities with potentially sparser or lower-quality open data. We trained this model as a proof-of-concept for minimally-congested times of day, but the same concept can be extended with similar training data for any other time of day if congestion data exist. Future research should also expand the kinds of training and validation data. Although used by billions of people around the world as the public's \enquote{best available} source of travel time prediction, Google travel time remains a black box based on users' GPS data. Future work can use other empirical travel time data to further train and validate the model.

@@

Travel time prediction is central to questions of urban accessibility, mode and route choice, and individuals' location decisions. True congested travel times vary drastically throughout the day and require large volumes of real-time or proprietary data, as well as complex algorithms, to model. These technical challenges and costs present a hurdle for many urban planners, who often turn to simpler models with lower and often free data and computing requirements. These traditional naïve methods may be easy to implement, but they produce wildly inaccurate predictions.

@@

>>>>

% !TeX program = pdflatex
% Travel Time Prediction from Sparse Open Data
% Authors: Geoff Boeing and Yuquan Zhou
% Web: https://geoffboeing.com/
% Repo: https://github.com/gboeing/travel-time-prediction

\newcommand{\myname}{Redacted for review}%Geoff Boeing and Yuquan Zhou}
\newcommand{\myemail}{boeing@usc.edu}
\newcommand{\myaffiliation}{Redacted for review}%University of Southern California}
\newcommand{\paperdate}{2025}
\newcommand{\papertitle}{Travel Time Prediction from Sparse Open Data}
\newcommand{\papercitation}{Boeing, G. and Y. Zhou. \paperdate. \papertitle. Under review at \emph{Journal Name}.}
\newcommand{\paperkeywords}{Urban Planning, Transportation, Data Science}

@@

\title{\papertitle}
\author[]{\myname}
\affil[]{\myaffiliation}
\date{}%\paperdate}

@@

Travel time prediction is central to transport geography and planning's accessibility analyses, sustainable transportation infrastructure provision, and active transportation interventions. However, calculating accurate travel times, especially for driving, requires either extensive technical capacity and bespoke data, or sources like the Google Maps API that quickly become prohibitively expensive to analyze thousands or millions of trips necessary for metropolitan-scale analyses. Such obstacles particularly challenge less-resourced researchers, practitioners, and community advocates. This article argues that a middle-ground is needed to provide reasonably accurate travel time predictions without extensive data or computing requirements. It introduces a free, open-source minimally-congested driving time prediction model with minimal cost, data, and computational requirements. It trains and tests this model using the Los Angeles, California urban area as a case study by calculating naïve travel times from open data then developing a random forest model to predict true travel times as a function of those naïve times plus open data on turns and traffic controls. Validation shows that this method offers a superior middle-ground technique that balances reasonable accuracy with minimal resource requirements.

@@

Travel time prediction underpins many transport planning processes and geographical research problems. It is key to understanding urban accessibility, transport mode and route choice, and location decisions \citep{zhang2015gradient,jenelius2013travel}. Predicting congested travel times (i.e., when traffic congestion impedes flows) requires large volumes of real-time geospatial data on traffic conditions \citep[e.g.,][]{hou2018network}. Such data are proprietary, expensive, and beyond the reach of many planning practitioners and scholars \citep{carrion2012value,wang2011estimating}. This is usually not a problem for well-resourced organizations with highly-skilled labor---however, GIScience technical expertise, complex data and software requirements, and high costs pose insurmountable obstacles for many urban planning researchers, practitioners, and community advocates \citep{giles2022creating}.

Travel time is fundamentally a function of distance and speed, which itself is a function of speed limits, traffic controls, turns, and congestion. Although common \enquote{naïve} travel time prediction methods (which minimize travel distance or street segment traversal time) provide easy solutions, they ignore the other components and thus usually under-predict real-world driving travel time \citep{ludwig2023traffic,salonen2013modelling,yiannakoulias2013estimating}. However, as we will argue, minimally-congested travel time prediction offers an important yet under-realized middle ground: much more accurate predictions than naïve travel times, yet much lower data and computing requirements than real-time traffic-aware congested travel time prediction.

To address the need for such a middle ground, this article contributes a novel accurate, easy-to-use, and free method of minimally-congested driving travel time prediction. Our results show a substantial improvement over traditional, naïve methods of travel time prediction common in the urban planning literature. As we shall show in this article, across the Los Angeles urban area, our average trip's travel time prediction differs from Google's \enquote{gold standard} prediction by only 0.38 seconds. In contrast, simple---but common in the literature---naïve travel time prediction minimizing max-speed edge traversal times under-predicts by 183.69 seconds on average.

The rest of this article is organized as follows. First it reviews the state of the art in travel time prediction and then considers the less-sophisticated methods that are standard in the current planning literature. Next it describes our proposed model, data sources, and validation techniques. Then it presents our findings, highlighting both the feasibility and quality of our predictions in relation to simple naïve predictions or difficult expensive predictions. Finally, it concludes with a discussion of implications for research and practice.

@@

Modern travel time prediction techniques tend to fall into one of two camps: (1) sophisticated, accurate, but difficult to execute; or (2) naïve, less-accurate, but simple to execute. The former is the state-of-the-art and has been the subject of much GIScience work particularly in the recent computer science and engineering literatures. The latter tends to appear often in urban planning research and practice. Here we summarize the recent advances and current standards in these fields.

@@

Travel time is essentially a function of distance, speed limits, traffic controls, turns, and traffic congestion. Naïve methods tend to predict it from just the first one or two of these components, whereas resource-intensive state-of-the-art methods predict it from all five. But, with open data, the first four components can now be freely acquired in many parts of the world---only congestion data remains universally proprietary, expensive, and inaccessible. Therefore, could there be a middle-ground between expensive, complicated, real-time predictions and overly-simple naïve predictions? Can we predict travel times from free open data on distance, speed limits, traffic controls, and turns? Such predictions may not reflect real-time congestion (which varies drastically throughout the day and week), but would accurately represent minimally-congested travel times.

This middle-ground method should meet three criteria. First, it should only use free software and free data, avoiding expensive API queries or bespoke GPS data collection. Second, it should be a simple and accessible method for urban planners without advanced technical skills. Third, it should provide a substantial improvement in accuracy and spatial resolution over naïve predictions such as minimizing Euclidean distance, network distance, street edge traversal time, and exsiting TAZ-to-TAZ travel times, and approach the accuracy of the state-of-art models. Meeting these three criteria, a middle ground would be inexpensive, easy-to-use, and reasonably accurate.

@@

This article proposes a novel open-source, reusable, generalizable method\endnote{Our open-source code is hosted at [REDACTED] for others to freely reuse.} to solve this problem. In short, it collects free open data plus a small amount of free data from the Google Maps API to train a local model to predict minimally-congested driving travel times with sparse data and simple computing hardware. \autoref{fig:workflow} shows a detailed workflow.

@@

After tuning, the four model training algorithms perform similarly across most---but not all---of the accuracy indicators (\autoref{tab:validation_results}). They fall roughly in a similar range for MAPE (7.9--9.0\%), MAE (72.0--80.0 seconds), MSE (12,155--13,570 square seconds), APR (0.99--1.01), and $R^2$ (0.93), demonstrating the robustness of our model to algorithmic particulars. However, the random forest and decision tree models exhibit much lower absolute systematic prediction bias ($\delta$ = 0.38 and 0.13 seconds, respectively) than the gradient boosting and AdaBoost models do (-19.2 and -9.8 seconds, respectively). In other words, all these models have comparable mean errors---but the average random forest and decision tree predictions are statistically insignificantly different from the average true travel time, whereas gradient boosting and AdaBoost significantly under-predict it. Therefore, we select the random forest model as our \enquote{final} travel time prediction model because it predicts better overall (than the decision tree model) and better avoids systematic under-prediction (than the gradient boosting and AdaBoost models).

This final model trains an ensemble of decision trees to predict unobserved response values by essentially dividing the data into boxes then making predictions based on the means of those boxes. Its optimized hyperparameterization uses 400 decision trees, random sampling with replacement in each decision tree, a maximum decision tree depth of 10, use of all available input explanatory features at each split, a requirement of at least two samples to split at a decision node, and the default settings that require at least one sample at a leaf node and apply equal sample weighting. Finally, we check this final model for overfitting by conducting another 5-fold cross-validation across the whole sample using the tuned hyperparameters. The resulting five MAE values (75.3, 74.5, 73.6, 73.2, 74.7) are all very similar, indicating that the model is not overfitted and its hyperparameters are not overly specific to the training set.

@@

\begin{table*}[tb!]
    \centering\small
    \caption{Out-of-sample prediction accuracy of our chosen model (random forest, bold) versus the initial naïve travel time model and the discarded alternative models, all validated against the corresponding true travel times. $n$ = 8,272 for each. The six accuracy indicators are (1) MAPE \%, (2) MAE in seconds, (3) MSE in square seconds, (4) difference-in-means ($\delta$) in seconds and its corresponding $t$-test's $p$-value, (5) APR, and (6) $R^2$. See methods for definitions.}\label{tab:validation_results}
    \tabularnums{\small
    \begin{tabular}{lrrrrrrr}
        \toprule
        Model               & MAPE  & MAE    & MSE      & $\delta$ & $p$  & APR  & $R^2$ \\
        \midrule
        Initial naïve       & 21.15 & 183.69 & 48214.06 &  -182.85 & <0.01 & 0.79 & 0.74  \\
        \textbf{Random forest}       &  \textbf{8.42} &  \textbf{75.32} & \textbf{12154.79} &     \textbf{0.38} &  \textbf{0.76} & \textbf{1.01} & \textbf{0.93}  \\
        Gradient boosting   &  7.86 &  71.99 & 12630.03 &   -19.20 & <0.01 & 0.99 & 0.93  \\
        Decision trees      &  9.00 &  80.04 & 13570.17 &     0.13 &  0.92 & 1.01 & 0.93  \\
        AdaBoost            &  8.20 &  74.04 & 12424.14 &    -9.76 & <0.01 & 1.00 & 0.93  \\
        \bottomrule
    \end{tabular}}
\end{table*}

Compared to the initial naïve travel time calculation, our travel time prediction model exhibits substantial out-of-sample improvement (\autoref{tab:validation_results}). Our model produces a MAPE of 8.4\%, in line with the \textasciitilde10\% MAPEs seen in the state-of-the-art travel time prediction literature, but without their extensive and expensive input data requirements. In comparison, our initial naïve travel time calculation produces a much worse MAPE of 21.2\%. Whereas the naïve travel time's MAE is 183.7 seconds, our model's predicted travel times' MAE is just 75.3 seconds---an improvement by a factor of 2.4. Similarly for MSE, our model improved by a factor of 4.0.

The $t$-test reveals significant differences between the initial naïve travel time calculations and the true travel times: the $\delta$ of -182.9 seconds corresponds to $p$ < 0.01. In other words, the naïve travel time under-predicts true travel time by over 3 minutes. However, the $t$-test reveals insignificant differences between our model's out-of-sample predictions and the true travel times: the $\delta$ of 0.38 seconds corresponds to a $p$ value of 0.76. That is, our prediction model over-predicts by less than half a second on average, which is statistically insignificantly different from zero. If we reconsider the aforementioned MAEs in light of these differences-in-means, we can see that the initial naïve travel time calculations significantly and consistently under-predict travel time, but our prediction model shows no such directional bias: its random error averages out between (smaller absolute) over-and under-prediction.

@@

Travel time is fundamentally a function of distance, speed limits, traffic controls, turns, and congestion. Naïve methods predict only from the first one or two of these components. Resource-intensive state-of-the-art methods predict it from all five. The planning literature shows that planners often use driving travel time prediction techniques far below the state-of-the-art due to cost, data, and technical capacity constraints. In this study, we propose that a better middle-ground travel time prediction method is feasible and identify three criteria for success: (1) it should use free software and data and avoid bespoke sensor data collection; (2) it should be easy and accessible to use; (3) it should offer better accuracy than naïve predictions (e.g., minimizing Euclidean distance, network distance, street edge traversal time) and better spatial resolution than zone-to-zone travel times.

@@

This middle ground is not necessarily for computer scientists, high-end labs with top-of-the-line computing hardware, or well-resourced agencies with extensive data budgets. Rather, it empowers the scholars and practitioners who need GIScience solutions the most---those working on the frontlines of urban planning and policymaking. The literature shows that these scholars and practitioners often rely on simple but inaccurate routing models---such as minimizing Euclidean distance, network distance, or edge traversal time---when they lack the technical expertise, computing equipment, or extensive data needed to implement the literature's cutting-edge models. However, this study demonstrates how these naïve methods systematically under-predict real-world travel times. Our model addresses this problem by incorporating sparse open data on traffic control elements and turns into the prediction to achieve an accuracy in line with state-of-the-art methods---but without their data needs or computing requirements. In other words, our model does not replace the state-of-the-art congested-traffic methods for those with extensive resources available to implement them, but rather offers far more accurate predictions for urban planners who would otherwise fall back on simpler naïve models. Inaccurate travel times skew empirical understandings and mislead planning interventions. Our model represents an important middle-ground contribution for urban planning scholarship and evidence-informed practitioner interventions.

As demonstrated, this model supports high-resolution point-to-point travel time predictions across a large metropolitan area. Our implementation focused on Los Angeles but is not inherently tied to it, and future work should emphasize its generalizability in cities with potentially sparser or lower-quality open data. We trained this model as a proof-of-concept for minimally-congested times of day, but the same concept can be extended with similar training data for any other time of day if congestion data exist. This is, of course, a big \enquote{if}---and the whole reason why minimally-congested predictions are valuable in the first place. Future research should also expand the kinds of training and validation data. Although used by billions of people around the world as the public's \enquote{best available} source of travel time prediction, Google travel time remains a black box based on users' GPS data. Future work can use other empirical travel time data to further train and validate the model.

@@

Travel time prediction is central to questions of urban accessibility, mode and route choice, and individuals' location decisions. True congested travel times vary drastically throughout the day and require large volumes of real-time or proprietary geospatial data, as well as complex algorithms, to model. These technical challenges and costs present a hurdle for many urban planners, who often turn to simpler models with lower and often free data and computing requirements. These traditional naïve methods may be easy to implement, but they produce wildly inaccurate predictions.

@@

\section*{Data Availability Statement}

The data that support the findings of this study are available at [OFFICIAL REPOSITORY URL REDACTED FOR ANONYMOUS PEER REVIEW.\@ REVIEWERS, PLEASE USE THIS ANONYMIZED URL FOR YOUR REVIEW INSTEAD:\@ https://figshare.com/s/0acd076b1a11286ff3e4].
%https://github.com/gboeing/travel-time-prediction

\section*{Conflict of Interest Statement}

The authors have no relevant financial or non-financial competing interests to report.
